Supplementary materials

### 2019-09-25 WholePlotTrajectories.Rmd
files

---
title: "Post-Disturbance Tree Community Trajectories in a Neotropical Forest"
author:
  - name: Ariane MIRABEL
    authsuperscript: 1*
  - name: Bruno Herault
    authsuperscript: 2
  - name: Eric Marcon
    authsuperscript: 1
affiliation:
  - affsuperscript: 1
    dptuniv: UMR EcoFoG, AgroParistech, CNRS, Cirad, INRA, Université des Antilles, Université de Guyane.
    address: >
      Campus Agronomique,
      97310 Kourou,
      France.
  - affsuperscript: 2
    dptuniv: INPHB, Institut National Polytechnique Félix Houphoüet-Boigny
    address: >
      Yamoussoukro,
      Ivory Coast.
corrauthor:
    email: ariane.mirabel@gmail.com
    url: https://github.com/ArianeMirabel
abstract: 
 In the current global change context, it is urgent to anticipate the fate of tropical forests.
 This means understanding tree community response to disturbance and the underlying processes.
 In that respect, we aim here to clarify taxonomic and functional post-disturbance  trajectories, and determine the scope of the Intermediate Disturbance Hypothesis (IDH) that remains debated in tropical forests.
 We analyzed community trajectories following a disturbance gradient in a Neotropical forest over 30 years. 
 We considered trajectories along time of community taxonomic and functional trajectories in terms of richness, evenness, composition, and redundancy.
 We based on the annual botanical inventories of 75 ha of a Neotropical forest and on large trait datasets comprising seven leaf, stem, and life-history traits. 
 We identified a decoupling between taxonomic composition, differing among communities, and functional composition, remaining similar and convergent. 
 The taxonomic diversity followed humped-shaped trajectories depending on the disturbance intensity, which validated the IDH (Intermediate Disturbance Hypothesis).
 The functional diversity trajectories,however, were homogeneous among all plots and dismissed the IDH.
 We explained this decoupling by the variations in community functional redundancy that mitigated the functional impact of disturbance. 
 Although consistent, the recovery of community composition, diversity, and redundancy remained unachieved after 30 years. These results acknowledged the need of decades-long cycles without disturbance to ensure community complete recovery, and questioned community resilience after repeated disturbances.
keywords: [Community Ecology, Disturbance Trajectories, Intermediate Disturbance Hypothesis, Mid-term Resilience, Neotropical Forests, Taxonomic and Functional Biodiversity]
preamble: >
  \usepackage{textcomp}
  \usepackage{tabu}
  \usepackage{caption}
  \captionsetup{justification = justified}
  \renewenvironment{table}{\begin{table*}}{\end{table*}\ignorespacesafterend}
  \hyphenation{bio-di-ver-si-ty sap-lings re-or-gan-i-za-tion post-dis-tur-bance dis-tur-bance}
bibliography: references.bib
lang: english
keywordlabel: Keywords
corrauthorlabel: Corresponding author
output:
  bookdown::pdf_book:
    base_format: EcoFoG::article
    keep_tex: TRUE
  bookdown::word_document2: default
  bookdown::gitbook: default
---

@@

```{r Tab1, echo=FALSE,as.is=T}
Intervention <- data.frame(
  Treatment=c("Control", "T1, low", "T2, intermediate", "T3, high"),
  Timber=c("-", 
           "DBH $\\geq$ 50 cm, commercially exploited species, $\\approx$ 10   $trees.ha^{-1}$", 
           "DBH $\\geq$ 50 cm, commercially exploited species, $\\approx$ 10  $trees.ha^{-1}$",
           "DBH $\\geq$ 50 cm, commercially exploited species, $\\approx$ 10  $trees.ha^{-1}$"),
  Thinning=c("-", 
             "-", 
             "DBH $\\geq$ 40 cm, non-commercially exploited species, $\\approx$ 30   $trees.ha^{-1}$", 
             "DBH $\\geq$ 50 cm, non-commercially exploited species, $\\approx$ 15  $trees.ha^{-1}$  "),
  Fuelwood=c("-", 
             "-", 
             "-", 
             "40 cm $\\leq$ DBH $\\leq$ 50 cm, non-commercially exploited species,\\ $\\approx$ 15 $trees.ha^{-1}$"),
  AGBLost=c(0, "$[12-33]$", "$[33-56]$", "$[35-56]$")
)
names(Intervention)[5] <- "\\%AGB lost"
knitr::kable(Intervention, caption="Intervention table, summary of the disturbance intensity for the 4 plot treatments in Paracou. Treatment intensities are defined by the minimum logging DBH (Diameter at Breast Height), the type of logged species (commercially exploited or not), the density of logged trees, and the total AGB (Above Ground Biomass) lost after treatment.", longtable = FALSE, booktabs = TRUE, escape = FALSE,format="latex") %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = TRUE)
```

@@

See Supplementary Materials -Fig. S1 and @Aubry-Kientz2013 for the detailed methodology.

@@

In control plots, Community Weighted Means (CWM) of functional traits remained stable in time.  
In disturbed plots, they mostly followed unimodal trajectories, either stabilizing or returning towards
their initial values, to the exception of leaf chlorophyll content, which continued to increase 30 years after disturbance for 4 out of 6 highly disturbed plots. Maximum height at adult stage (*Hmax*), leaf toughness, and wood specific gravity (*WSG*) decreased in time and then slightly increased, but remained significantly lower than their initial value (Fig. \ref{fig:CWM}). 
Bark thickness and specific leaf area (*SLA*) both increased along time. 
Bark thickness remained substantially high after 30 years, and *SLA* had almost recovered to its initial value. Whatever the functional traits, the maximum
difference to initial value was highly correlated to the disturbance intensity.
Positive correlations were observed for leaf thickness, chlorophyll content, SLA and bark thickness ($\rho_{Spearman}^{Leaf thickness}=0.76$, $\rho_{Spearman}^{Chlorophyll content}=0.60$,
$\rho_{Spearman}^{SLA}=0.93$, $\rho_{Spearman}^{Bark thickness}=0.71$). 
Negative correlation was observed for leaf toughness, WSG, and Hmax ($\rho_{Spearman}^{Leaf toughness}=-0.53$,
$\rho_{Spearman}^{WSG}=-0.75$, $\rho_{Spearman}^{Hmax}=-0.40$)
The proportions of the three lightest seed mass classes increased in all disturbed plots. 
After 30 years the proportion of lightest seed mass class decreased while it stabilized for the two other lightest seed mass classes (Supp. Mat. - Fig. S2).

>>>>

---
title: "Post-Disturbance Tree Community Trajectories in a Neotropical Forest"
author:
  - name: Ariane MIRABEL
    authsuperscript: 1*
  - name: Bruno Herault
    authsuperscript: 2
  - name: Eric Marcon
    authsuperscript: 1
affiliation:
  - affsuperscript: 1
    dptuniv: UMR EcoFoG, AgroParistech, CNRS, Cirad, INRA, Université des Antilles, Université de Guyane.
    address: >
      Campus Agronomique,
      97310 Kourou,
      France.
  - affsuperscript: 2
    dptuniv: INPHB, Institut National Polytechnique Félix Houphoüet-Boigny
    address: >
      Yamoussoukro,
      Ivory Coast.
corrauthor:
    email: ariane.mirabel@gmail.com
    url: https://github.com/ArianeMirabel
abstract: 
 In the current global change context, it is urgent to anticipate the fate of tropical forests.
 This means understanding tree community response to disturbance and the underlying processes.
 In that respect, we aim here to clarify taxonomic and functional post-disturbance  trajectories, and determine the scope of the Intermediate Disturbance Hypothesis (IDH) that remains debated in tropical forests.
 We analyzed community trajectories following a disturbance gradient in a Neotropical forest over 30 years. 
 We considered trajectories along time of community taxonomic and functional trajectories in terms of richness, evenness, composition, and redundancy.
 We based on the annual botanical inventories of 75 ha of a Neotropical forest and on large trait datasets comprising seven leaf, stem, and life-history traits. 
 We identified a decoupling between taxonomic composition, differing among communities, and functional composition, remaining similar and convergent. 
 The taxonomic diversity followed humped-shaped trajectories depending on the disturbance intensity, which validated the IDH (Intermediate Disturbance Hypothesis).
 The functional diversity trajectories,however, were homogeneous among all plots and dismissed the IDH.
 We explained this decoupling by the variations in community functional redundancy that mitigated the functional impact of disturbance. 
 Although consistent, the recovery of community composition, diversity, and redundancy remained unachieved after 30 years. These results acknowledged the need of decades-long cycles without disturbance to ensure community complete recovery, and questioned community resilience after repeated disturbances.
keywords: [Community Ecology, Disturbance Trajectories, Intermediate Disturbance Hypothesis, Mid-term Resilience, Neotropical Forests, Taxonomic and Functional Biodiversity]
preamble: >
  \usepackage{textcomp}
  \usepackage{tabu}
  \usepackage{caption}
  \captionsetup{justification = justified}
  \renewenvironment{table}{\begin{table*}}{\end{table*}\ignorespacesafterend}
  \hyphenation{bio-di-ver-si-ty sap-lings re-or-gan-i-za-tion post-dis-tur-bance dis-tur-bance}
  \usepackage[left]{lineno}
  \linenumbers
bibliography: references.bib
lang: english
keywordlabel: Keywords
corrauthorlabel: Corresponding author
output:
  bookdown::word_document2: default
  bookdown::pdf_book:
    base_format: EcoFoG::article
    keep_tex: TRUE
  bookdown::gitbook: default
---

@@

```{r Tab1, echo=FALSE,as.is=T}
Intervention <- data.frame(
  Treatment=c("Control", "T1, low", "T2, intermediate", "T3, high"),
  Timber=c("-", 
           "DBH $\\geq$ 50 cm, commercially exploited species, <br/> $\\approx$ 10   $trees.ha^{-1}$", 
           "DBH $\\geq$ 50 cm, commercially exploited species, $\\approx$ 10  $trees.ha^{-1}$",
           "DBH $\\geq$ 50 cm, commercially exploited species, $\\approx$ 10  $trees.ha^{-1}$"),
  Thinning=c("-", 
             "-", 
             "DBH $\\geq$ 40 cm, non-commercially exploited species, $\\approx$ 30   $trees.ha^{-1}$", 
             "DBH $\\geq$ 50 cm, non-commercially exploited species, $\\approx$ 15  $trees.ha^{-1}$  "),
  Fuelwood=c("-", 
             "-", 
             "-", 
             "40 cm $\\leq$ DBH $\\leq$ 50 cm, non-commercially exploited species,\\ $\\approx$ 15 $trees.ha^{-1}$"),
  AGBLost=c(0, "$[12-33]$", "$[33-56]$", "$[35-56]$")
)
names(Intervention)[5] <- "\\%AGB lost"
knitr::kable(Intervention, caption="Intervention table, summary of the disturbance intensity for the 4 plot treatments in Paracou. Treatment intensities are defined by the minimum logging DBH (Diameter at Breast Height), the type of logged species (commercially exploited or not), the density of logged trees, and the total AGB (Above Ground Biomass) lost after treatment.", longtable = FALSE, booktabs = TRUE, escape = FALSE,format="latex") %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = TRUE)
```

@@

See Supplementary Materials -Fig. A.1 and @Aubry-Kientz2013 for the detailed methodology.

@@

In control plots, Community Weighted Means (CWM) of functional traits remained stable in time.  
In disturbed plots, they mostly followed unimodal trajectories, either stabilizing or returning towards
their initial values, to the exception of leaf chlorophyll content, which continued to increase 30 years after disturbance for 4 out of 6 highly disturbed plots. Maximum height at adult stage (*Hmax*), leaf toughness, and wood specific gravity (*WSG*) decreased in time and then slightly increased, but remained significantly lower than their initial value (Fig. \ref{fig:CWM}). 
Bark thickness and specific leaf area (*SLA*) both increased along time. 
Bark thickness remained substantially high after 30 years, and *SLA* had almost recovered to its initial value. Whatever the functional traits, the maximum
difference to initial value was highly correlated to the disturbance intensity.
Positive correlations were observed for leaf thickness, chlorophyll content, SLA and bark thickness ($\rho_{Spearman}^{Leaf thickness}=0.76$, $\rho_{Spearman}^{Chlorophyll content}=0.60$,
$\rho_{Spearman}^{SLA}=0.93$, $\rho_{Spearman}^{Bark thickness}=0.71$). 
Negative correlation was observed for leaf toughness, WSG, and Hmax ($\rho_{Spearman}^{Leaf toughness}=-0.53$,
$\rho_{Spearman}^{WSG}=-0.75$, $\rho_{Spearman}^{Hmax}=-0.40$)
The proportions of the three lightest seed mass classes increased in all disturbed plots. 
After 30 years the proportion of lightest seed mass class decreased while it stabilized for the two other lightest seed mass classes (Supp. Mat. - Fig. A.2).

### 2019-09-25 WholePlotTrajectories_SuppMat.Rmd
files

---
title: "Supplementary materials"
author:
  - name: Ariane MIRABEL
    authsuperscript: 1*
  - name: Eric MARCON
    authsuperscript: 1
  - name: Bruno HERAULT
    authsuperscript: 2
affiliation:
  - affsuperscript: 1
    dptuniv: UMR EcoFoG, AgroParistech, CNRS, Cirad, INRA, Université des Antilles, Université de Guyane.
    address: >
      Campus Agronomique,
      97310 Kourou,
      France.
  - affsuperscript: 2
    dptuniv: INPHB (Institut National Polytechnique Félix Houphoüet Boigny) 
    address: Yamoussoukro, Ivory Coast
corrauthor:
    email: ariane.mirabel@ecofog.gf
    url: https://github.com/ArianeMirabel
preamble: >
  \usepackage{textcomp}
  \DeclareUnicodeCharacter{B0}{\textdegree}
  \usepackage{tabu}
  \renewenvironment{table}{\begin{table*}}{\end{table*}\ignorespacesafterend}
  \hyphenation{bio-di-ver-si-ty sap-lings}
lang: english
output: 
  bookdown::pdf_book:
    base_format: EcoFoG::article
    keep_tex: TRUE
  bookdown::gitbook: default
  bookdown::word_document2: default
---

@@

@@

@@

```{r FigBayesianScheme,out.width='100%',fig.cap="Scheme of the Bayesian process propagating the taxonomic uncertainty",echo=F, fig.env='figure',fig.pos='h'}
knitr::include_graphics("ExternalFig/BayesianSheme.png")
```

@@

```{r FigSeedMassClasses, echo=FALSE, out.width='80%', fig.asp=2/5,fig.cap="Trajectories of seed mass classes proportions over 30 years after disturbance. Colors are treatments: green (control), yellow (T1), orange (T2), red (T3).", fig.env='figure',fig.pos='h'}
load("DB/Smass_ClassProp")

@@

@@

```{r PCA_graphs, out.width = '100%', echo=FALSE, fig.cap="Summarizing plots of the functional PCA analysis defining communities functional space. (ref:Bolda) Plot of the samples from the BRIDGE functional database in the two dimensional main plan. (ref:Boldb) bow plot of the PCA eignevalues in percentage. (ref:Boldc) Plot of the 7 functional traits sed in the PCA analysis (Leaf and stem traits, and maximum species height).", fig.env='figure',fig.pos='h'}
load("DB/FunctionalPCA")

>>>>

---
title: "Supplementary materials"
author:
  - name: Ariane MIRABEL
    authsuperscript: 1*
  - name: Eric MARCON
    authsuperscript: 1
  - name: Bruno HERAULT
    authsuperscript: 2
affiliation:
  - affsuperscript: 1
    dptuniv: UMR EcoFoG, AgroParistech, CNRS, Cirad, INRA, Université des Antilles, Université de Guyane.
    address: >
      Campus Agronomique,
      97310 Kourou,
      France.
  - affsuperscript: 2
    dptuniv: INPHB (Institut National Polytechnique Félix Houphoüet Boigny) 
    address: Yamoussoukro, Ivory Coast
corrauthor:
    email: ariane.mirabel@ecofog.gf
    url: https://github.com/ArianeMirabel
preamble: >
  \usepackage{textcomp}
  \DeclareUnicodeCharacter{B0}{\textdegree}
  \usepackage{tabu}
  \renewenvironment{table}{\begin{table*}}{\end{table*}\ignorespacesafterend}
  \hyphenation{bio-di-ver-si-ty sap-lings}
  \captionsetup{labelformat=empty}
lang: english
output: 
  bookdown::pdf_book:
    base_format: EcoFoG::article
    keep_tex: TRUE
  bookdown::gitbook: default
  bookdown::word_document2: default
---

@@

```{r, include=F}
library(captioner)
figures <- captioner(prefix = "Figure S", auto_space = FALSE)

figures("FigBayesianScheme", "Scheme of the Bayesian process propagating the taxonomic uncertainty", display=FALSE)
figures("FigSeedMassClasses", "Trajectories of seed mass classes proportions over 30 years after disturbance. Colors are treatments: green (control), yellow (T1), orange (T2), red (T3).", display=FALSE)
figures("PCAgraphs", "Summarizing plots of the functional PCA analysis defining communities functional space. (ref:Bolda) Plot of the samples from the BRIDGE functional database in the two dimensional main plan. (ref:Boldb) bow plot of the PCA eignevalues in percentage. (ref:Boldc) Plot of the 7 functional traits sed in the PCA analysis (Leaf and stem traits, and maximum species height).", display=FALSE)
```

@@

#Supplementary Material {-}

\pagenumbering{arabic}

@@

```{r FigBayesianScheme,out.width='100%',fig.cap=figures("FigBayesianScheme"),echo=F, fig.env='figure',fig.pos='h'}
knitr::include_graphics("ExternalFig/BayesianSheme.png")
```

\pagebreak

@@

```{r FigSeedMassClasses, echo=FALSE, out.width='80%', fig.asp=2/5,fig.cap=figures("FigSeedMassClasses"), fig.env='figure',fig.pos='h'}
load("DB/Smass_ClassProp")

@@

\pagebreak

@@

```{r PCAgraphs, echo=FALSE, out.width = '100%', fig.cap=figures("PCAgraphs"), fig.env='figure',fig.pos='h'}
load("DB/FunctionalPCA")

### 2019-09-25 WholePlotTrajectories_SuppMat.tex
final text
files

% User-adder preamble
\usepackage{textcomp} \DeclareUnicodeCharacter{B0}{\textdegree}
\usepackage{tabu}
\renewenvironment{table}{\begin{table*}}{\end{table*}\ignorespacesafterend}
\hyphenation{bio-di-ver-si-ty sap-lings}

@@

\Authors{
Ariane MIRABEL\textsuperscript{1*}\\ Eric MARCON\textsuperscript{1}\\ Bruno HERAULT\textsuperscript{2}
} % Authors
\affiliation{
\textsuperscript{1}UMR EcoFoG, AgroParistech, CNRS, Cirad, INRA, Université des Antilles,
Université de Guyane.\\ \hspace{1em} Campus Agronomique, 97310 Kourou, France.\\\textsuperscript{2}INPHB (Institut National Ploytechnique Félix Houphoüet Boigny)\\ \hspace{1em} Yamoussoukro, Ivory Coast
}
\affiliation{*\textbf{}: ariane.mirabel@ecofog.gf, https://github.com/ArianeMirabel} % Corresponding author

@@

The following information are available for this article:

@@

\caption{Scheme of the Bayesian process propagating the taxonomic uncertainty}\label{fig:FigBayesianScheme}
\end{figure}

@@

\caption{Trajectories of seed mass classes proportions over 30 years after disturbance. Colors are treatments: green (control), yellow (T1), orange (T2), red (T3).}\label{fig:FigSeedMassClasses}
\end{figure}

@@

{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_SuppMat_files/figure-latex/PCA_graphs-1}

@@

\caption{Summarizing plots of the functional PCA analysis defining communities functional space. \textbf{(a)} Plot of the samples from the BRIDGE functional database in the two dimensional main plan. \textbf{(b)} bow plot of the PCA eignevalues in percentage. \textbf{(c)} Plot of the 7 functional traits sed in the PCA analysis (Leaf and stem traits, and maximum species height).}(\#fig:PCA_graphs)
\end{figure}

>>>>

% User-adder preamble
\usepackage{textcomp} \DeclareUnicodeCharacter{B0}{\textdegree}
\usepackage{tabu}
\renewenvironment{table}{\begin{table*}}{\end{table*}\ignorespacesafterend}
\hyphenation{bio-di-ver-si-ty sap-lings}
\captionsetup{labelformat=empty}

@@

\Authors{
Ariane MIRABEL\textsuperscript{1*}\\ Eric MARCON\textsuperscript{1}\\ Bruno HERAULT\textsuperscript{2}
} % Authors
\affiliation{
\textsuperscript{1}UMR EcoFoG, AgroParistech, CNRS, Cirad, INRA, Université des Antilles,
Université de Guyane.\\ \hspace{1em} Campus Agronomique, 97310 Kourou, France.\\\textsuperscript{2}INPHB (Institut National Polytechnique Félix Houphoüet Boigny)\\ \hspace{1em} Yamoussoukro, Ivory Coast
}
\affiliation{*\textbf{}: ariane.mirabel@ecofog.gf, https://github.com/ArianeMirabel} % Corresponding author

@@

\section*{Supplementary Material}\label{supplementary-material}
\addcontentsline{toc}{section}{Supplementary Material}

\pagenumbering{arabic}

@@

\caption{Figure S1: Scheme of the Bayesian process propagating the taxonomic uncertainty}\label{fig:FigBayesianScheme}
\end{figure}

\pagebreak

@@

\caption{Figure S2: Trajectories of seed mass classes proportions over 30 years after disturbance. Colors are treatments: green (control), yellow (T1), orange (T2), red (T3).}\label{fig:FigSeedMassClasses}
\end{figure}

\pagebreak

@@

{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_SuppMat_files/figure-latex/PCAgraphs-1}

@@

\caption{Figure S3: Summarizing plots of the functional PCA analysis defining communities functional space. \textbf{(a)} Plot of the samples from the BRIDGE functional database in the two dimensional main plan. \textbf{(b)} bow plot of the PCA eignevalues in percentage. \textbf{(c)} Plot of the 7 functional traits sed in the PCA analysis (Leaf and stem traits, and maximum species height).}\label{fig:PCAgraphs}
\end{figure}
